\chapter{Mitigating Trade-Offs with Commonalities 
    \pgsize{35 p.}
}
\label{chap:improvement}

\mnote{Topologies and Properties}
We have identified in the previous chapter that the topology of the graph induced by the metamodels and transformations of a transformation network directly influences several of its quality properties, such as functional correctness and completeness as well as maintainability in terms of modularity and reusability.
The extreme topologies of complete graphs and trees imply extremes in the optimization or degradation of these properties, which induces a trade-off between these properties by means of the topology.

\mnote{Benefits of trees}
In \autoref{part:correctness}, we have focused on achieving correctness for networks of arbitrary topology, thus in general not inducing a tree but any graph topology that can be extended to a complete graph, which inherently optimizes reusability and completeness but requires high effort for achieving completeness.
On the contrary, a tree structure, although not that easy to achieve, provides inherent correctness guarantees while reducing reusability and completeness (see \autoref{chap:classification:topologies:effects}).
In this chapter, we discuss how a network having a tree topology can be constructed by introducing additional metamodels, such that correctness is still given but reusability and completeness is improved.

\mnote{Common concepts}
The idea of adding metamodels is not only a conceptual necessity to improve quality properties but also motivated by practical benefits.
Since consistency relations define how common information is represented in several metamodels redundantly, we propose to represent this common information explicitly by means of additional metamodels.
Then, only the manifestation of this information in the models to keep consistent has to be defined rather than an implicit encoding of common information in the consistency relations.
These manifestation relations can, of course, again be represented by transformations.
This way of specifying consistency with explicit metamodels representing common information can inherently lead to a transformation network with a tree topology.

%\todo{Motivation for Comnmonalities idea: it is common in other domains (Albers) to model the overlaps of models for preserving consistency (check slides of SFB and reference their publications)}

\mnote{Subordinate contributions}
This chapter constitutes our contribution \autoref{contrib:quality:improvement}, which consists of four subordinate contributions: a discussion of how common information can be represented explicitly in dedicated metamodels and under which conditions this is reasonable; a proposal of the \commonalities approach to construct such metamodels and transformations for describing the manifestations of common information in the original metamodels; a discussion of the expected benefits of the approach, especially in terms of mitigating trade-offs between quality properties; and finally an outlook to processes of applying the approach and of combining it with other transformations. It answers the following research question:

\researchquestionrepeat{rq:quality:topology}

\mnote{Benefits of contributions}
The insights in this chapter support transformation developers in constructing networks of correct, complete, and reusable transformations.
It gives a different view on consistency and the possibilities to describe it besides consistency relations, which we expect to improve comprehensibility due to common concepts being represented explicitly rather than encoding them in consistency relations implicitly.
The proposed construction approach for transformation networks inherently improves several quality properties by reducing the effort to achieve correctness of transformation networks as discussed in \autoref{part:correctness} and mitigating necessary trade-offs.
It especially improves reusability and completeness in comparison to an ordinary construction of a network having a tree topology.

\mnote{Publication of contributions}
The initial idea for the contributions in this chapter has already been published~\owncite{klare2018docsym} as well as the proposed \commonalities approach with its expected benefits~\owncite{klare2019models}.
The approach along with a language that supports it, which we present in the subsequent chapter, has originally been developed in the Bachelor's thesis of \textowncite{gleitze2017a}, which was supervised by the author of this thesis.


\input{sections/6_quality/621_improvement-concepts.tex}
\input{sections/6_quality/622_improvement-commonalities.tex}
\input{sections/6_quality/623_improvement-benefits.tex}
\input{sections/6_quality/624_improvement-application.tex}
\input{sections/6_quality/625_improvement-summary.tex}
