\chapter{Designing a Language for Expressing Commonalities 
    \pgsize{25 p.}
}
\label{chap:language}

\mnote{\Commonalities approach}
In the previous chapter, we have introduced the \commonalities approach, which defines a methodology for constructing transformation networks by means of auxiliary, so called \conceptmetamodels.
These \conceptmetamodels contain the commonalities of the metamodels whose instances are to be kept consistent, denoted as \concretemetamodels, as explicit entities, rather than encoding them implicitly in transformations between the metamodels to be kept consistent.
We have argued why this construction approach foster achieving a specific tree topology of the transformation network.
Such a topology improves correctness and reusability of the resulting transformation networks, which are contradictory properties when constructing networks only of transformations between the \concretemetamodels, at least if a specific tree topology of the network is achieved.

\mnote{\Commonalities language}
Although the construction methodology of the \commonalities approach itself provides significant benefits and is thus a distinct and independently usable contribution on its own, the construction can be further supported with an appropriate language.
While the approach requires the specification of \conceptmetamodels, as well as transformations realizing the manifestation relations between the metamodels, a language can combine these specification by integrating the definition of manifestation with those of the \commonalities.
This improves conciseness and locality of the related information to be defined.
While those improvements only foster usability, but provide no conceptual benefits, a language can also ensure the achievement of an appropriate tree topology.
This can either be achieved by restricting expressiveness to achieve it by construction or by defining constructs that are a analyzable.

\mnote{Subordinate contributions}
In this chapter, we discuss the design of such a language.
We focus on design options for such a language and given an overview of the process and artifacts involved in such a language.
We also depict a concrete language, for which we have developed the prototypical \commonalities language, with a focus on the relevant elements, their relations and their operationalization.
Although we also provide a prototypical realization of such a language, this chapter does not focus on the specifics of that language, but rather the concepts behind it.
It constitutes our contribution \autoref{contrib:quality:language}, which consists of three subordinate contributions: a discussion of design options and the resulting process and artifacts for such a language; a depiction of the structure of a concrete realization of such a language with a description of its semantics and operationalization into transformations; and finally a discussion of benefits that we can expect from such a language.
It answers the following research question:

\researchquestionrepeat{rq:quality:language}

\mnote{Benefits of contributions}
The insights in this chapter first give guidelines for developers of tools for construction transformation networks.
It especially clarifies the available design space for tools supporting the \commonalities approach.
In addition, the chapter makes concrete proposals for how to develop such a language, which elements it has to contain and how it can be operationalized.
Finally, it even provides an actual realization of such a language, which can be readily used with the \vitruv framework (see \autoref{chap:foundations:multiview:vitruv}).

\mnote{Publication of contributions}
An overview of the prototypical realization of the \commonalities language and relevant design options along with a proof-of-concept has already been published~\owncite{klare2019models}.
An initial prototype of the language was developed in the Bachelor's thesis of \textowncite{gleitze2017a} and extended for a case study evaluation in the Master's thesis of \textowncite{hennig2020ma}.
Since we focus on the concepts and design options for such a language in this thesis, we refer to the work of \textowncite{gleitze2017a} for details about the realization and capabilities of the \commonalities language.


\input{sections/6_quality/631_language-design_options.tex}
\input{sections/6_quality/632_language-elements.tex}
\input{sections/6_quality/633_language-benefits.tex}
\input{sections/6_quality/634_language-summary.tex}
