A dissertation is not written by one person alone.
So many people are involved in that process in many different ways that I cannot thank and address all of them by name.
Therefore, a first big thanks goes to everyone who supported me in this adventure.

My deepest thanks go to Ralf.
He gave me the opportunity to work as a researcher and to write a dissertation, and he inspired me in countless discussions. It was a pleasure to work on visions and research directions at eye level and to see that my opinions are valued.
But most importantly, he always had a friendly ear and gave support whenever and for whatever reason that was necessary. 

I thank Colin for all the great and inspiring meetings we had and all the valuable and constructive feedback he gave me. Having all these talks with someone from outside our chair clearly helped be to think outside of the box and better frame my ideas.

I am grateful to everyone at SDQ who worked with me, discussed ideas with me, or simply made the time enjoyable.
Special thanks go the \vitruvius guys Max, Michael, and Erik. Without them, I would not even have considered writing a dissertation.
And without all the great students that supported me in realizing my ideas, it would have taken at least twice the time to finish this thesis. 
Therefore, particular thanks go to Aurélien, Joshua, Timur, and Torsten.
I am especially thankful for all the friends I made along the way. Representative for them, I thank Max for being a friend for years now, Dominik and Stephan for all the great help and the fun we had, and Sandro for many sympathetic talks and the guidance in baking sourdough bread.

I thank my family and my friends for all the love, the sympathy, and every possible way of support, which allowed me to frame my life so that I could place this thesis into it.
I am so lucky and infinitely grateful to be surrounded by such people, who take me as I am and stand by me for so many year.
These final thanks go to my parents, to Kati, to Marius, and to Mirco.
