\chapter{Correctness in Transformation Networks
    \pgsize{45 p.}
}
\label{chap:correctness}

%\todo{Discuss hippocraticness (although it may not be necessary). Bereits implizit in der consistency notion eingeführt. Auch auf Idempotenz der Funktion verweisen.}

\mnote{Chapter goals}
In this chapter, we will first discuss a rather informal notion of consistency and its preservation. It is supposed to describe the different dimensions in which consistency and its preservation can be considered to then discuss how \emph{correctness} can be reasonably defined.
After identifying the correctness notion that is relevant in the context of our work, we define a suitable formal notion of consistency.
We formally define correctness of different artifacts relevant for that notion of consistency.
Finally, we present a refined notion of consistency, which we do not require for the initial overview, but which we will later use for several detailed considerations.

\mnote{Contributions}
This chapter thus constitutes our contribution \autoref{contrib:correctness:notion}, which is composed of four subordinate contributions: a discussion of consistency notions; a discussion and determination of correctness notions for consistency specifications; a formalization of a relevant correctness notion; and finally a refinement of our consistency notion for later detailed considerations.
It answers the following research question:

\researchquestionrepeat{rq:correctness:notions}

\mnote{Publication of contributions}
Parts of the contributions in this chapter have already been published in \owncite{klare2020Vitruv-JSS}, \owncite{klare2020compatibility-report} and \owncite{klare2018docsym}.
In \owncite{klare2020Vitruv-JSS}, we have used a simplified version of the formalization that we introduce in this chapter and especially identified the challenge of orchestration, which is central for the formalization of transformation networks.
\owncite{klare2020compatibility-report} discussed compatibility and introduced the fine-grained consistency notion, which was required for detailed statements about compatibility.
In \owncite{klare2018docsym}, we especially motivated and informally derived the correctness notion that we formalize in the following and give an overview of the goal regarding correctness of transformation networks.

\input{sections/3_correctness/3110_correctness-consistency_notion.tex}
\input{sections/3_correctness/3120_correctness-correctness_notion.tex}
\input{sections/3_correctness/3130_correctness-formalization.tex}
\input{sections/3_correctness/3140_correctness-finegrained.tex}
\input{sections/3_correctness/3150_correctness-summary.tex}
