\part{Prologue}
\input{sections/1_prologue/110_introduction}

\begin{researchgoal}
	A transformation developer shall know about all necessary properties of a transformation network to achieve its correct execution and he or she shall be provided with techniques to guarantee these properties by construction whenever possible and otherwise techniques that allow him or her to check them.
\end{researchgoal}

\section{Research Questions}

\subsection*{Properties of Transformation Networks}
\begin{enumerate}[label=\itshape RQ \arabic*.]
	% Issue classification
	\item Which issues can occur when independently developed \acp{BX} are combined to a network?
	\begin{nestedenum}
		\item Which failures can occur, when \acp{BX} are combined to a network?
		\item What mistakes can be made that lead to failures?
		\item How can these mistakes be categorized regarding conceptual levels in the specification process for \acp{BX}?
	\end{nestedenum}
	\item How are properties of transformation networks affected by the topology?
	\begin{nestedenum}
		\item Which properties are relevant when defining networks of \acp{BX}?
		\item Which topologies of network exist and how do they affect that properties?
	\end{nestedenum}
\end{enumerate}

\subsubsection*{Evaluation}

Keine dezidierte Evaluation, lediglich Argumentation


\subsection*{Building Correct Transformation Networks} % Correctness of Transformation Networks?
\begin{enumerate}[label=\itshape RQ \arabic*.]
	\setcounter{enumi}{2}
	% Contradiction analysis
	\item How can transformations be analyzed regarding contradictions in specified constraints?
	\begin{nestedenum}
		\item What is an appropriate formalism for describing transformations that can be analyzed regarding potential contradictions?
		\item Which kinds of contradictions can be detected by analyzing transformations following a specific formalism?
	\end{nestedenum}
	% Issue avoidance by construction
	\item How can interoperability of independently developed \acp{BX} be achieved by construction?
	\begin{nestedenum}
		\item Which kinds of mistakes can be avoided by construction of the individual \acp{BX}?
		\item How can we prove that those mistakes and only those mistakes can be avoided by construction?
		\item How can each of these mistakes be avoided by a transformation developer during independent development of a single \acp{BX}?
	\end{nestedenum}
	% Orchestration
	\item What is an appropriate strategy for orchestrating independently developed \acp{BX} to perform a fixed-point iteration?
	\begin{nestedenum}
		\item Which strategies for orchestrating a network of \acp{BX} exists and what are their properties?
		\item How should those properties be weighted and which of the strategies should be chosen for orchestration?
	\end{nestedenum}
\end{enumerate}

\subsubsection*{Evaluation}

\gqm{Functionality}{The analysis can be used to find contradictions in specifications}
{Does the analysis find contradictions if they exist?}
{Recall: Ratio of true positives to true positives + false negatives}
\qm{Does the analysis find contradictions although they do not exist?}
{Precision: Ratio of true positives to true+false positives}
\qm{Does the analysis find non-contradictions although they exist?}
{Ratio of false negatives to false+true negatives}

\gqm{Functionality}{The techniques to avoid mistakes by construction actually avoid interoperabililty issues}
{Are the identified failures that can occur complete?}
{Ratio of number of identified failures to total number of failures}
\qm{Are the relations of identified mistakes to identified failures correct?}
{Ratio of failures resolved by fixing the identified mistake to all failures}
\qm{Does the application of avoidance techniques lead to interoperable transformations?}
{Ratio of changes that are propagated correctly to those that are not propagated correctly}

\gqm{Applicability}{The techniques can be applied independently to single transformations}
{Are there cases in which information about other transformations are necessary to solve issues?}
{Ratio of number of fixes that require information about other transformation to total number of fixes with user interactions\\
Ratio of number of fixes that require information about other transformation to total number of fixes without user interactions}



\subsection*{Improving Non-Functional Properties of Transformation Networks}

\begin{enumerate}[label=\itshape RQ \arabic*.]
	\setcounter{enumi}{5}
	\item How can a topology of transformation be build that optimizes non-functional properties of transformation networks?
	\begin{nestedenum}
	    \item How can transformation contradictions be avoided by language design? %uniqueness of consistency specification and consistency among themselves achieved by language design?
	    \item How can modularity be achieved in a way such that an arbitrary set of \metamodels for which consistency is specified can be used in an actual project?
	\end{nestedenum}
	% Building tree topologies
	\item How should a language specific for multi-model consistency be defined that supports a non-functional property-optimizing topology definition?
	\begin{nestedenum}
		\item What are the design decision for such a language?
	\end{nestedenum}
\end{enumerate}

\subsubsection*{Evaluation}

\gqm{Functionality}{Concept and language can achieve consistency between several models}{How many model changes in a case study can be properly kept consistent?}{Ratio of successfull test cases}

\gqm{Practicality}{The assumption of defining a tree of \commonalities is achievable in practice}{Is the definition of cross-tree relations necessary in a case study?}{Number of cross-tree relations in a case study compared to number of relations}

\gqm{Practicality/Benefit}{A specific language improves conciseness of consistency specifications}{How much more concise is the specification for a case study compared to a definition with direct transformations?}{Number of SLOC with \commonalities compared to number of SLOC with \reactions for same case study}

Diskussion: Erreichen der Modularität auch evaluieren? Ist per Konstruktion gegeben, könnte man aber natürlich auch noch auswerten (bringt aber nichts).


\input{sections/1_prologue/120_foundations}
