\chapter{Conclusions \pgsize{5 p.}}

%% FROM ICMT
\subsection{Interoperability Issues}
Issues that can arise from the combination of independently developed \acp{BX} to networks have not been systematically investigated yet.
In this paper, we therefore categorized failures that can occur when executing faulty networks of \acp{BX}.
Additionally, we structured the process of specifying consistency into three levels: the global level, the modularization level and the operationalization level.
These levels carry the danger for different kinds of mistakes, which we categorized and related to potential failures they can result in.
We found that each of the levels is prone to different types of mistakes, and that each type of failure is specific for one category of mistake.
This enables developers to easily identify the kind of mistake they made when recognizing a failure.
Additionally, the systematic knowledge about potential mistakes, failures, and their relations makes it possible to further develop techniques to avoid them.
We have discussed two general avoidance strategies at the modularization and operationalization level in this paper.
In future work, we will especially investigate how far and under which assumptions \acp{BX} can be analyzed regarding contradictions at the modularization level when they are combined.


%% FROM VOSE
\subsection{Quality Properties by Commonalities}

In this paper, we proposed the \commonalities approach, which allows to make common concepts of different \metamodels explicit.
The central idea of the presented approach 
%Its central idea 
is to define \emph{\conceptmetamodels} that represent the common concepts, i.e., the \commonalities of two or more existing \metamodels, and to define the relation of those \metamodels to the \conceptmetamodels.
\Conceptmetamodels can be hierarchically composed to enable the separate definition and combination of independent concepts.
We discussed different options for designing a language that supports the specification of such \commonalities and their relations, 
as well as for operationalizing such a specification to executable transformations.
We outlined a language for the \commonalities approach and explained %our selection of the aforementioned options. %
which of the aforementioned options we chose, and why.
%We have implemented the language as a proof-of-concept and applied it to simple scenarios.
Finally, we have applied an implementation of that language to simple scenarios as a proof-of-concept.
The results indicate the feasibility of applying the \commonalities approach and implementing an appropriate language.

The expected benefit of our approach is a better understandability of relations between \metamodels compared to their implicit encoding in transformations.
Additionally, we argued why our approach improves the essential properties \emph{compatibility} and \emph{modularity}, which usually contradict each other in other approaches to keep multiple models consistent, like networks of transformations that define the direct relations between \metamodels.
In ongoing work, we extend the capabilities of the language to perform a comprehensive evaluation of the functionality of the approach as well as its applicability to more sophisticated case studies.
We will validate our claim of improved understandability in a controlled experiment.
Nevertheless, the initial results of our proof-of-concept are a promising indicator for the applicability of the \commonalities approach.
